Probabilistic Programming

### 2014-04-21 paper/paper.tex
Add screenshots and discussion of vim autocomplete

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}

@@

We work with programs at the level of abstract syntax tree (AST), using the
syntax tree format of Mozilla's Parser API \cite{parser api}. We use
esprima \cite{esprima}, an ECMAScript parser, to transform program source code into AST,
and escodegen \cite{escodegen} (ECMAScript code generator) to transform ASTs into program source
code.

@@

We encode the structure of an AST as a probabilistic model. The model is a data
structure that maps node paths to probability distributions of node types or
values. A node path is a list of node types and relationships that encodes the
unique position of a node in a program's AST. The probability distribution for a
node path is the set of possible node types or values that would appear in the
syntax tree following the given path, and the respective probabilities of their
occurance following the given path.

@@

The vim autocomplete script

The guided IDE

@@

\begin{figure}[h!]
	\caption{model-sample.json}
	\centering
	A model generated from the program ``var x = 34"
\end{figure}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}

@@

We work with programs at the level of abstract syntax tree (AST), using the
syntax tree format of Mozilla's Parser API \cite{parser api}. We use
esprima \cite{esprima}, an ECMAScript parser, to transform program source code into an AST,
and escodegen \cite{escodegen} (ECMAScript code generator) to transform ASTs into program source
code.

@@

We encode the structure of an AST as a probabilistic model. The model is a data
structure that maps node paths to probability distributions of node types or
values. A node path is a list of node types and relationships that encodes the
unique position of a node in a program's AST. The probability distribution for a
node path is the set of possible node types or values that would appear in the
syntax tree following the given path, and the respective probabilities of their
occurrence following the given path.

The model is implemented as a JSON object where the path is indexed into the
object by successively indexing by each string of the path.

See figure~\ref{fig:sample-model} in the appendix for a model for the program
``var x = 34;".

@@

\subsubsection{vim autocomplete script}

We integrated our model into the text editor vim to offer autocompletion
suggestions of programs upon the user's request, using vim's
\texttt{completefunc} feature. When the user is editing a JS file in vim with
this feature, they may press a keyboard command to get a list of program
snippets to choose from. The snippets are generated in the context of the
program path at cursor's position. The user may pick one of the snippet
suggestions to have it inserted into their code at their cursor.

The autocomplete snippets are generated as follows:
\begin{enumerate}
	\item Parse the contents of the current file into an AST.
	\item Find the node path corresponding to the cursor by traversing the AST.
		%This represents the program context at which the code snippets should
		%be generated.
	\item Generate programs starting from the node path of the cursor.
	\item Present the programs to the user as code completion snippets.
\end{enumerate}

The vim autocomplete feature is limited to generating full programs on a single
line, and requires explicitly triggering the \texttt{completefunc} feature for
code snippets.

\begin{figure}[h]
	\centering
	\begin{tabular}{c c}

		\begin{subfigure}[h]{6.25cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim1}
			\caption{Inserting a for loop} \label{fig:vim1}
		\end{subfigure}

		\begin{subfigure}[h]{6cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim2}
			\caption{Inserting an object property} \label{fig:vim2}
		\end{subfigure}
	\end{tabular}

	\caption{Screenshots of vim using autocomplete}
\end{figure}

\subsubsection{Guided IDE}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.90\textwidth]{screenshot}
  \caption{Screenshot of the web program editor GUI} \label{fig:screenshot}
\end{figure}

@@

\begin{figure}[h!]
	\caption{model-sample.json}
	\label{fig:sample-model}
	\centering
	A model generated from the program ``var x = 34;"
\end{figure}
